---
output: html_document
---

## How-to
***

#### Exponential Smoothing

For an explanation of the ideas of exponential smoothing, I can't recommend more the second chapter (free) of [Forecasting with Exponential Smoothing: A State Space Approach](http://www.exponentialsmoothing.net/toc) by Rob J. Hyndman, Anne B. Koehler, J. Keith Ord, and Ralph D. Snyder


#### Example

We will now illustrate a particular case of estimating an exponential smoothing problem, taken from the previously mentioned book.

Our data will be the inbuilt dataset in *R*, *USAccDeaths*, a dataset consisting of Accidental Deaths in the US 1973-1978. We will fit an "optimal" exponential smoothing model by using an automatic inbuilt algorithm in the *ets* *R* package and forecast 12 months of data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(forecast)
library(ggplot2)
library(gridExtra)
library(zoo)

data("USAccDeaths")
Dates <- as.Date(time(USAccDeaths))
DeathsDF <- data.frame(Dates, USAccDeaths)
fit <- ets(USAccDeaths)

fcast <- forecast(fit, 12)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)

# I credit the GitHub user fernandotenorio for this function.
new.plot <-function(forec.obj, data.color = 'black', fit.color = 'blue', forec.color = 'red',
                           lower.fill = 'orange', upper.fill = 'darkorange', format.date = F)
{
    serie.orig = forec.obj$x
    serie.fit = forec.obj$fitted
    pi.strings = paste(forec.obj$level, '%', sep = '')
    
     if(format.date)
        dates = as.Date(time(serie.orig))
    else
        dates = time(serie.orig)
    
    serie.df = data.frame(date = dates, serie.orig = serie.orig, serie.fit = serie.fit)
    
    forec.M = cbind(forec.obj$mean, forec.obj$lower[, 1:2], forec.obj$upper[, 1:2])
    forec.df = as.data.frame(forec.M)
    colnames(forec.df) = c('forec.val', 'l0', 'l1', 'u0', 'u1')
    
    if(format.date)
        forec.df$date = as.Date(time(forec.obj$mean))
    else
        forec.df$date = time(forec.obj$mean)
    
    p = ggplot() + theme_minimal()  + xlab("") + ylab("USAccDeaths") +
        geom_line(aes(date, serie.orig, colour = 'data'), data = serie.df) + 
        geom_line(aes(date, serie.fit, colour = 'fit'), data = serie.df) + 
        scale_y_continuous() +
        geom_ribbon(aes(x = date, ymin = l0, ymax = u0, fill = 'lower'), data = forec.df, alpha = I(0.4)) + 
        geom_ribbon(aes(x = date, ymin = l1, ymax = u1, fill = 'upper'), data = forec.df, alpha = I(0.3)) + 
        geom_line(aes(date, forec.val, colour = 'forecast'), data = forec.df) + 
        scale_color_manual('Series', values=c('data' = data.color, 'fit' = fit.color, 'forecast' = forec.color), labels = c("Original Data", "Fitted Values", "Forecast") ) + 
        scale_fill_manual('Confidence Interval', values=c('lower' = lower.fill, 'upper' = upper.fill), labels = pi.strings)
    
    if (format.date)
        p = p + scale_x_date()
    
    p
}
new.plot(fcast)

summary(fcast)
```

In order to obtain this plot set:

- Error Type: Aditive
- Trend Type: None
- Season Type: Aditive

Or, if you're feeling a little bit lazy, just click the "Fit and Forecast" button as the default values and data are set to deliver our example plot.

***
#### Prohibited combinations

Let the first element stand for the "Error Type" specification, the second element for the "Trend type" specification and the third element for the "Season Type" specification. Let *N* denote "None", *A* "Aditive", and *M* "Multiplicative". The following are non allowed ETS model specifications:

- ANM
- AMN
- AMA
- AAM
- AMM
- MMA